How to Analyze Build Logs---A~Comparative Study of Chunk~Retrieval~Techniques

### 2020-04-03 tse-paper/content.tex
small fixes in study sections

\subsection{Program Synthesis by Example}
\label{sec:rw-prose}
% % \section{Information Extraction and Retrieval Techniques} The
% % techniques we investigate are based on existing methods of
% % Programming by Example.%, information extraction and information
% % retrieval.
% This section presents different Programming by Example
% % resources surrounding the PROSE library.
% We explain its generic
% % program synthesis algorithm and how PROSE synthesizes text
% % extraction programs, the foundation of PBE\@.

@@

\section{Chunk Retrieval Techniques}
\label{sec:techniques}
This section introduces key concepts of the three chunk retrieval
techniques we study in this article, namely program synthesis by
example (PBE), common text similarity (CTS), and keyword search (KWS).
We use these techniques to extract text chunks from build logs that
represent a specific, targeted information.

@@

% TODO Caro: I find this figure still very confusing :'(, due to a lack
of structure.
If you want me to, I can give it a go?
\begin{figure*}[tb]
	\centering
	\includegraphics[width=\textwidth, trim={0.4cm 8.4cm 1.2cm 0.3cm},
	clip]{img/study.pdf}
	\caption{Study design of our technique comparison study.}
	\label{fig:study}
\end{figure*}

To investigate which techqnique is best suited to retrieve chunks from
CI build logs we evaluate them on the
\emph{LogChunks}~\cite{brandt2020logchunks} data set.
This section
describes our study design and which metrics we measure to answer our
research questions.
In the presentation of the results, we first focus
on each of the three techniques and later compare them against each
other.

@@

For the comparison, we evaluate the three chunk retrieval techniques
PBE, CTS and KWS, described in Sections~\ref{sec:expl-pbe},
\ref{sec:expl-ts} and \ref{sec:expl-skws}.
Random Line Retrieval
(RLR), explained in Section~\ref{sec:expl-rlr}, acts as a baseline for
the comparison.
% We run four techniques on the examples from \emph{LogChunks}.

\noindent
\textbf{Training and Test Set}
We use \emph{LogChunks} as the data set for our study.
For each of the
80 repositories in it, \emph{LogChunks} contains 10 build logs,
manually labeled with the \emph{chunk} describing the
reason the build failed.
Additionally, it contains (1) keywords useful
to identify the substring in the log  and (2) which structural category
the substring belongs to.
% TODO Moritz: You said: Add details how key words were extracted.
% Add example?
% how the keywords were put in logchunks does not fit here, that is
% maybe answered 2 paragraphs above already?
% how KWS uses the keywords is already described in the section
% introducing the technique => I would scrap everything but the first
% sentence directly above here :)

@@

\noindent % TODO: adapt if we change the research questions
\textbf{RQ 2.1: Size of Training and Test Set}
To analyze how many examples the chunk retrieval techniques need to
perform best, we evaluate the techniques with different training set
sizes.
We train each technique with one to five examples from each of
the repositories within \emph{LogChunks}.
The size of the test set is
one.

\noindent
\textbf{RQ 2.2: Recording Structural Categories}
To determine how structurally similar the examples for the chunk
retrieval techniques need to be, we record the structural categories
of the examples in the training and test sets.

\noindent
\textbf{RQ2.3: Accuracy Metrics}
To measure the accuracy of the retrieved chunks we save the output
lines of the chunk retrieval run on the input of the test example
($\mathit{RetrievedLines}$).
As oracle in our evaluation, we save the
desired lines from the output of the test example
($\mathit{DesiredLines}$).

@@

\vspace{0.2cm}
\begin{itemize}[leftmargin=0.4cm] \itemsep1em
	\item $|\mbox{True positives}| = \mathit{DesiredLines} \cap
	\mathit{RetrievedLines}$ \vspace{0.2cm}\\
	True positives are lines that appear both in the output of a
	tool as well as what \textit{LogChunks} contains for each
	example.

@@

	\item $\mbox{Recall} =
	\dfrac{|\mathit{TruePositives}|}{|\mathit{DesiredLines}|}$
	\vspace{0.2cm} \\
	Recall of a chunk retrieval describes which proportion of the
	desired lines were retrieved.
	\item $\mbox{F$_{1}$-score} = 2 \cdot \dfrac{\mathit{Precision}
	\cdot \mathit{Recall}}{\mathit{Precision} + \mathit{Recall}}$
	\vspace{0.2cm}\\
	In addition, we calculate the F$_{1}$-score, the harmonic mean
	of precision and recall.
We prefer F$_{1}$ to other aggregate
	measures such as accuracy because for a
	``needle-in-the-haystack'' scenario, we do not want to bloat
	our results by correctly not finding lots of irrelevant log
	lines.
	\item Successful retrieval = $\mathit{true}\ \mathit{iff}\
	\mathit{Recall} = 1, \mathit{false\ otherwise}$  \vspace{0.2cm} \\
	We define a successful retrieval as one where all desired
	lines were
	extracted, therefore when recall is one.
\end{itemize}

% Recall and precision of CTS and KWS vary with the number of lines
% selected for retrieval.
% We evaluate the effect of varying the number
% of extracted lines by multiplying the average number of lines
% present in the training examples with a \emph{retrieval size factor}
% from 0.5 to 2.5 in steps of 0.5.

\section{Results}
This section presents the results for PBE, CTS and KWS separately.
Afterwards we compare the three techniques with each other and RLR as
baseline.

\lstset{
  language=,
  morekeywords={Test, Output, Desired},
	basicstyle=\ttfamily\scriptsize,
  postbreak=\mbox{\textcolor{cyan}{$\hookrightarrow$}\space},
  showspaces=false,
  showstringspaces=false,
  keywordstyle=\textbf,
	frame=single,
  extendedchars=false,
  texcl=false
}
\begin{figure}[!t]
  \centering
  \begin{lstlisting}[breaklines=true]
Test Output:
Error: Invalid CSS after "2.3em": expected expression (e.g.
1px, bold), was ";"
	on line 86 of sass/components/dropdown.sass
Desired Test Output:
Error: Invalid CSS after "2.3em": expected expression (e.g.
1px, bold), was ";"
	on line 86 of sass/components/dropdown.sass
	from line 5 of sass/components/_all.sass
	from line 6 of bulma.sass
  \end{lstlisting}
  \caption{Example for an unsuccessful retrieval (PBE retrieved only
  two of the four targeted lines).}
  \label{lst:pbe-unsuccessful}
\end{figure}

\subsection{Common Text Similarity (CTS)}
\label{sec:r:cts}
\begin{figure*}
\centering
    \textbf{Common Text Similarity (CTS)}\par\medskip
\begin{subfigure}[t]{\columnwidth}
		\centering
		\includegraphics[width=\columnwidth,
		clip]{img/big-study/recall-precision-examplecount-CTS.pdf}
		\caption{Precision, recall and F$_{1}$-score for an
		increasing training set size.}
		\label{fig:recall-precision-examplecount-CTS}

\end{subfigure}\hspace{\fill}
\begin{subfigure}[t]{\columnwidth}
		\centering
				\includegraphics[width=\columnwidth,
				clip]{img/big-study/recall-precision-categorycount-CTS.pdf}
		\caption{Precision, recall and F$_{1}$-score for
		an increasing number of structural categories in the
		training and test sets.}
		\label{fig:recall-precision-categorycount-CTS}
\end{subfigure}
\caption{Results of chunk retrieval with Common Text Similarity (CTS)}
\end{figure*}

@@

\subsection{Keyword Search (KWS)}
\label{sec:r:kws}
\begin{figure*}
\centering
    \textbf{Keyword Search (KWS)}\par\medskip
\begin{subfigure}[b]{\columnwidth}
		\centering
		\includegraphics[width=\columnwidth,
		clip]{img/big-study/recall-precision-examplecount-KWS.pdf}
		\caption{Precision, recall and F$_{1}$-score for an
		increasing training set size.}
		\label{fig:recall-precision-examplecount-KWS}
\end{subfigure}\hspace{\fill}
\begin{subfigure}[b]{\columnwidth}
		\centering
		\includegraphics[width=\columnwidth,
		clip]{img/big-study/recall-precision-categorycount-KWS.pdf}
		\caption{Precision, recall and F$_{1}$-score for
		an increasing number of structural categories in the
		training and test sets.}
		\label{fig:recall-precision-categorycount-KWS}
\end{subfigure}
\caption{Results of chunk retrieval with Keyword Search (KWS)}
\end{figure*}

@@

% Figure~\ref{fig:contextsizefactor-precision-recall-KWS} shows the
% effect of the retrieval size factor on precision, recall and
% F$_{1}$-score of chunk retrieval runs with KWS\@.
%The precision is
% 19\% when retrieving half of the expected number of lines.
%On
% average 9\% of the lines in the build log are retrieved then.
%When
% 2.5 times the expected number of lines are retrieved, the precision
% decreases to 10\% and a quarter of the lines in the build log are
% retrieved on average.
%The recall ranges from 58\% to 75\% and the
% F$_{1}$-score shows a constant decrease from 29\% to 17\%.

% \begin{figure}[!t]
%		\centering
%		\includegraphics[width=\columnwidth,
%clip]{img/big-study/contextsizefactor-precision-recall-KWS.pdf}
%		\caption{Precision, recall and F$_{1}$-score of chunk
%retrieval with KWS compared to retrieval size factor.}
%		\label{fig:contextsizefactor-precision-recall-KWS}
% \end{figure}

\subsection{Program Synthesis by Example (PBE)}
\label{sec:r:pbe}
\begin{figure*}
\centering
    \textbf{Program Synthesis by Example (PBE)}\par\medskip
\begin{subfigure}[b]{\columnwidth}
		\centering
		\includegraphics[width=\columnwidth,
		clip]{img/big-study/recall-precision-examplecount-sythesisworked-PBE.pdf}
				\caption{Precision, recall and
				F$_{1}$-score when PBE could synthesize
				a consistent program compared with the
				size of the training set.}
		\label{fig:recall-precision-examplecount-sythesisworked-PBE}
\end{subfigure}\hspace{\fill}
\begin{subfigure}[b]{\columnwidth}
		\centering
		\includegraphics[width=\columnwidth,
		clip]{img/big-study/failure-reason-categorycount-PBE.pdf}
				\caption{Increasing number of structural
				categories in the training and test sets.}
		\label{fig:failure-reason-categorycount-PBE}
\end{subfigure}
\caption{Results of chunk retrieval with  Program Synthesis by Example
(PBE)}
\end{figure*}

@@

In our evaluation, we compare against a baseline of randomly
picking lines from the build log.
Its results follow intuitive
expectations:
The precision ranges between 8\% and 5\%, the recall between 6\% and
8\%.
The size of the training set has no visible impact on either.
An greater number of structural categories within the training and
test sets decreased the precision of RLR from 7\% to 0\%, the recall
increases from 6\%, for one structural category present, to 11\%, for
three sturctural categories present and also drops to 0\% for more
structural categories.
Graphs of the results of the chunk retrieval with RLR from our study
are included in our replication
package~\cite{brandt2020chunk-replication}.

\subsection{Comparison of All Techniques}
In this section, we aggregate the results from
\Cref{sec:r:kws,sec:r:pbe,sec:r:cts,sec:r:rlr} and put them next to it
each other.

@@

\begin{figure*}
\centering
    \textbf{PBE, CTS, KWS, and RLR}\par\medskip
\begin{subfigure}[b]{\columnwidth}
		\centering
				\includegraphics[width=\columnwidth,
				clip]{img/big-study/recall-precision-singlecategory-all.pdf}
		\caption{Precision, recall and F$_{1}$-score of all
		techniques when training examples are in \emph{one}
		structural category.}
		\label{fig:recall-precision-singlecategory-all}
\end{subfigure}\hspace{\fill}
\begin{subfigure}[b]{\columnwidth}
		\centering
				\centering
		\includegraphics[width=\columnwidth,
		clip]{img/big-study/recall-precision-multicategory-all.pdf}
		\caption{Precision, recall and F$_{1}$-score of all
		techniques when training examples are in \emph{more than
		one} structural categories.}
		\label{fig:recall-precision-multicategory-all}
\end{subfigure}
\caption{Comparison of all chunk retrieval techniques, split by structural
category count.}
\end{figure*}

>>>>

\subsection{Program Synthesis by Example}
\label{sec:rw-prose}
% % \section){Information Extraction and Retrieval Techniques} The
% % techniques we investigate are based on existing methods of
% % Programming by Example.%, information extraction and information
% % retrieval.
% This section presents different Programming by Example
% % resources surrounding the PROSE library.
% We explain its generic
% % program synthesis algorithm and how PROSE synthesizes text
% % extraction programs, the foundation of PBE\@.

@@

\section{Chunk Retrieval Techniques}
\label{sec:techniques}
% TODO: rework! extend / relate to further techniques
This section introduces key concepts of the three chunk retrieval
techniques we study in this article, namely program synthesis by
example (PBE), common text similarity (CTS), and keyword search (KWS).
We use these techniques to extract text chunks from build logs that
represent a specific, targeted information.

@@

% TODO Caro: redo
\begin{figure*}[tb]
	\centering
	\includegraphics[width=\textwidth, trim={0.4cm 8.4cm 1.2cm 0.3cm},
	clip]{img/study.pdf}
	\caption{Study design of our technique comparison study.}
	\label{fig:study}
\end{figure*}

To investigate which technique is best suited to retrieve chunks from
CI build logs we evaluate the three chunk retrieval techniques
PBE, CTS and KWS, described in Sections~\ref{sec:expl-pbe},
\ref{sec:expl-ts} and \ref{sec:expl-skws}.
Random Line Retrieval
(RLR), explained in Section~\ref{sec:expl-rlr}, acts as a baseline for
the comparison.
This section
describes our study design and which metrics we measure to answer our
research questions.
In the presentation of the results, we first focus
on each of the three techniques and later compare them against each
other.

@@

% We run four techniques on the examples from \emph{LogChunks}.

\noindent
\textbf{Training and Test Set}
% TODO Moritz: You said: Add details how key words were extracted.
% Add example?
% how the keywords were put in logchunks does not fit here, that is
% maybe answered 2 paragraphs above already?
% how KWS uses the keywords is already described in the section
% introducing the technique => I would scrap everything but the first
% sentence directly above here :)

@@

\noindent % TODO: adapt if we change the research questions
\textbf{RQ 2.1: Size of Training and Test Set}
We evaluate the techniques with different training set
sizes, to analyze how many examples the chunk retrieval techniques need to
perform best.
We train each technique with one to five examples from each of
the repositories within \emph{LogChunks}.
The size of the test set is one.

\noindent
\textbf{RQ 2.2: Recording Structural Categories}
We record the structural categories
of the examples in the training and test sets, to determine how
structurally similar the examples for the chunk
retrieval techniques need to be.

\noindent
\textbf{RQ2.3: Accuracy Metrics}
we save the output lines of the chunk retrieval run on the input
of the test example ($\mathit{RetrievedLines}$), to measure the
accuracy of the retrieved chunks 
As oracle in our evaluation, we save the
desired lines from the output of the test example
($\mathit{DesiredLines}$).

@@

\vspace{0.2cm}
\begin{itemize}[leftmargin=0.4cm] \itemsep1em
	\item $|\mbox{True\ Positives}| = \mathit{DesiredLines} \cap
	\mathit{RetrievedLines}$ \vspace{0.2cm}\\
	True positives are lines that appear both in the output of a
	tool as well as what \textit{LogChunks} contains for the
  test example.

@@

	\item $\mbox{Recall} =
	\dfrac{|\mathit{True\ Positives}|}{|\mathit{DesiredLines}|}$
	\vspace{0.2cm} \\
	Recall of a chunk retrieval describes which proportion of the
	desired lines were retrieved.
	\item $\mbox{F$_{1}$-score} = 2 \cdot \dfrac{\mathit{Precision}
	\cdot \mathit{Recall}}{\mathit{Precision} + \mathit{Recall}}$
	\vspace{0.2cm}\\
	In addition, we calculate the F$_{1}$-score, the harmonic mean
	of precision and recall.
We prefer F$_{1}$ to other aggregate
	measures such as accuracy because for a
	``needle-in-the-haystack'' scenario, we do not want to bloat
	our results by correctly not finding lots of irrelevant log
	lines.
	\item Successful retrieval = $\mathit{true}\ \mathit{iff}\
	\mathit{Recall} = 1, \mathit{false\ otherwise}$  \vspace{0.2cm} \\
	We define a successful retrieval as one where all desired
	lines were
	extracted, therefore when recall is one.
\end{itemize}

\section{Results}
This section presents the separate results for each chunk retrieval
technique. We start with CTS, KWS and RLR, as their results are easiest to
understand.
Afterwards we compare the three techniques with each other and RLR as
baseline.

\lstset{
  language=,
  morekeywords={Test, Output, Desired},
  keywordstyle=\textbf,
	frame=single
}
\begin{figure}[!t]
  \centering
  \begin{lstlisting}[breaklines=true]
Test Output:
Error: Invalid CSS after "2.3em": expected expression (e.g.
1px, bold), was ";"
	on line 86 of sass/components/dropdown.sass
Desired Test Output:
Error: Invalid CSS after "2.3em": expected expression (e.g.
1px, bold), was ";"
	on line 86 of sass/components/dropdown.sass
	from line 5 of sass/components/_all.sass
	from line 6 of bulma.sass
  \end{lstlisting}
  \caption{Example for an unsuccessful retrieval (PBE retrieved only
  two of the four targeted lines).}
  \label{lst:pbe-unsuccessful}
\end{figure}

\subsection{Common Text Similarity (CTS)}
\label{sec:r:cts}
\begin{figure*}
\centering
    \textbf{Common Text Similarity (CTS)}\par\medskip
\begin{subfigure}[t]{\columnwidth}
		\centering
		\includegraphics[width=\columnwidth,
		clip]{img/big-study/recall-precision-examplecount-CTS.pdf}
		\caption{Compared with the training set size.}
		\label{fig:recall-precision-examplecount-CTS}

\end{subfigure}\hspace{\fill}
\begin{subfigure}[t]{\columnwidth}
		\centering
				\includegraphics[width=\columnwidth,
				clip]{img/big-study/recall-precision-categorycount-CTS.pdf}
		\caption{Compared with the number of structural categories in the
		training and test sets.}
		\label{fig:recall-precision-categorycount-CTS}
\end{subfigure}
\caption{Precision, recall and F$_{1}$-score of chunk
retrieval with Common Text Similarity (CTS)}
\end{figure*}

@@

\subsection{Keyword Search (KWS)}
\label{sec:r:kws}
\begin{figure*}
\centering
    \textbf{Keyword Search (KWS)}\par\medskip
\begin{subfigure}[b]{\columnwidth}
		\centering
		\includegraphics[width=\columnwidth,
		clip]{img/big-study/recall-precision-examplecount-KWS.pdf}
		\caption{Compared with the training set size.}
		\label{fig:recall-precision-examplecount-KWS}
\end{subfigure}\hspace{\fill}
\begin{subfigure}[b]{\columnwidth}
		\centering
		\includegraphics[width=\columnwidth,
		clip]{img/big-study/recall-precision-categorycount-KWS.pdf}
		\caption{Compared with the number of structural categories in the
		training and test sets.}
		\label{fig:recall-precision-categorycount-KWS}
\end{subfigure}
\caption{Precision, recall and F$_{1}$-score of chunk retrieval with
Keyword Search (KWS)}
\end{figure*}

@@

In our evaluation, we compare against a baseline of randomly
picking lines from the build log.
Its results follow intuitive
expectations:
The precision ranges between 8\% and 5\%, the recall between 6\% and
8\%.
The size of the training set has no visible impact on either.
An larger number of structural categories within the training and
test sets decreased the precision of RLR from 7\% to 0\%, the recall
increases from 6\%, for one structural category present, to 11\%, for
three sturctural categories present and also drops to 0\% for more
structural categories.
Graphs of the results of the chunk retrieval with RLR from our study
are included in our replication
package~\cite{brandt2020chunk-replication}.

\subsection{Program Synthesis by Example (PBE)}
\label{sec:r:pbe}
\begin{figure*}
\centering
    \textbf{Program Synthesis by Example (PBE)}\par\medskip
\begin{subfigure}[b]{\columnwidth}
		\centering
		\includegraphics[width=\columnwidth,
		clip]{img/big-study/recall-precision-examplecount-sythesisworked-PBE.pdf}
				\caption{Precision, recall and
				F$_{1}$-score when PBE could synthesize
				a consistent program compared with the
				size of the training set.}
		\label{fig:recall-precision-examplecount-sythesisworked-PBE}
\end{subfigure}\hspace{\fill}
\begin{subfigure}[b]{\columnwidth}
		\centering
		\includegraphics[width=\columnwidth,
		clip]{img/big-study/failure-reason-categorycount-PBE.pdf}
				\caption{Successfulness of retrieval compared with the number of
        structural categories in the training and test sets.}
		\label{fig:failure-reason-categorycount-PBE}
\end{subfigure}
\caption{Results of chunk retrieval with  Program Synthesis by Example
(PBE)}
\end{figure*}

@@

\subsection{Comparison of All Techniques}
Next, we aggregate the results from
\Cref{sec:r:kws,sec:r:pbe,sec:r:cts,sec:r:rlr}.

@@

\begin{figure*}
\centering
    \textbf{PBE, CTS, KWS, and RLR}\par\medskip
\begin{subfigure}[b]{\columnwidth}
		\centering
				\includegraphics[width=\columnwidth,
				clip]{img/big-study/recall-precision-singlecategory-all.pdf}
		\caption{Training and test examples are in \emph{one}
		structural category.}
		\label{fig:recall-precision-singlecategory-all}
\end{subfigure}\hspace{\fill}
\begin{subfigure}[b]{\columnwidth}
		\centering
				\centering
		\includegraphics[width=\columnwidth,
		clip]{img/big-study/recall-precision-multicategory-all.pdf}
		\caption{Training and test examples are in \emph{more than
		one} structural categories.}
		\label{fig:recall-precision-multicategory-all}
\end{subfigure}
\caption{Precision, recall and F$_{1}$-score of all
techniques compared, split by structural category count.}
\end{figure*}
